```{r}
current_path = rstudioapi::getActiveDocumentContext()$path
setwd(dirname(current_path))
source("anovatab.R")
```

```{r}
#Done on exercise_class4
pain=read.table("pain.txt",header=T)
pain
```
###fit One way analysis of variance (ANOVA)  
```{r}
treatF <- factor(pain$treat)
fit_pain=lm(Relief~treatF,data=pain)
summary(fit_pain)
```
- Stage I Analysis: As p-value is < 0.05 on F-statistic, we reject the null hypothesis (that there is not relationship between treat and Relief) 
- Remember that $\hat{\beta}_1$ is actually estimate of $t_2 - t_1$ and $\hat{\beta}_2$ is $t_3 - t_1$

```{r}
anovatab(fit_pain)
```
  
#### Creating dummy variables by myself:
```{r}

d1=d2=d3=d4=rep(0,nrow(pain))
d1[pain$treat==1]=1
d2[pain$treat==2]=1
d3[pain$treat==3]=1
d4[pain$treat==4]=1
pain=cbind(pain,d1,d2,d3,d4)


```

#### Fitting model with dummy
```{r}
fit_d_pain <- lm(Relief~d2+d3+d4, data=pain)
summary(fit_d_pain)

```
  
#### Fitting model with dummy with all variables
```{r}
fit_d2_pain <- lm(Relief~d1+d2+d3+d4, data=pain)
summary(fit_d2_pain)

```

#### Return to regular fitting 
```{r}
summary(fit_pain)
```
  


```{r}
library(multcomp)
```
  

### Apply General Linear Hyphotesis 
#### compare t2 vs t3 -> $t2 - t3$
```{r}
L=cbind(0,1,-1,0)
glh1=glht(fit_pain,linfct=L)
summary(glh1,test=Ftest())
```

- P-value is 0.62 so there is no evidence that t2 is different from t3 (codeine vs acumpuncture)

#### compare t3 vs t4 -> $t3 - t4$ (acupuncture vs acupuncture+codeine)
```{r}
L=cbind(0,0,1,-1)
glh1=glht(fit_pain,linfct=L)
summary(glh1,test=Ftest())

```
- P-value is small so there is evidence that the combined treatment t4 is more effective than t3

#### compare t2 vs t4 -> $t2 - t4$ (codeine vs acupuncture+codeine)
```{r}
L=cbind(0,1,0,-1)
glh1=glht(fit_pain,linfct=L)
summary(glh1,test=Ftest())

```
- P-value is small so there is evidence that the combined treatment t4 is more effective than t2 as well


#### compare t1 vs t4 -> $t1 - t4$ (nothing vs acupuncture+codeine)
```{r}
L=cbind(0,0,0,1)
glh1=glht(fit_pain,linfct=L)
summary(glh1,test=Ftest())

```