---
html_document:
  mathjax: http://cdn.mathjax.org/mathjax/latest/MathJax.js?config=TeX-AMS-MML_HTMLorMML
author: "Mauricio Reis | D21125621"
title: "Assignment 2"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

Complete the following project using R Markdown and ensure that your code and output are readable in the html file.

### The `veteran` dataset

```{r}
library(survival)
?veteran
```

#### Talking points

Hello, welcome to this video

-   We gonna use the Veteran dataset that represents data from a randomised trial of two treatments for lung cancer.\
-   Here in this analisys we gonna focus into two [[things:\\\\](things:\\){.uri}]([things:\\](things:\){.uri}){.uri}

1.  The relevance of having prior treatment for the survival of each patient (prior 0 to no prior and 10 to prior).\
2.  The effect of age of the patient for the survival

-   Also relevant for this video, we have time that represents the survival time for each patient and status that represents censoring status.\

```{r}
head(veteran)
```

### Questions

#### Talking points Kaplan Meier

-   We estimate and plotted the survival function for the two prior groups using Kaplan-Meier estimator.

-   We also found the probability of survival for each one of those prior groups for time bigger than 200.

1.  Estimate and plot $S(t)$ for the two `prior` groups using the Kaplan-Meier estimator and find $\Pr(T > 200)$ for an individual with `prior` = 0 and `prior` = 10. [**25 marks**]\

```{r}
#Create a Survival object using time and status to be the response and say it's explained by prior
SKM <- survfit(Surv(time, status) ~ prior, data = veteran)
summary(SKM)
```

```{r}
plot(SKM, 
     col = c("blue", "red"),
     lwd = 3,
     main = "S(t) for the two prior groups on veteran Dataset\n marking point of survival beyond t = 200")

legend("topright", 
       c("prior = 0", "prior = 10"), 
       lty = 1,
       lwd = 3,
       col = c("blue", "red"))
S_200_prior0 <- summary(SKM, times = 200)$surv[1]
S_200_prior10 <- summary(SKM, times = 200)$surv[2]
points(200, y = S_200_prior0, type = "p", col = "black", pch = 16)
points(200, y = S_200_prior10, type = "p", col = "black", pch = 16)

```

### Talking points

-   We can see here the survival function for prior 0 and 10 as well as where the point above 200 is located.\
-   This gives an impression that patient with previous treatment have a higher survival probability at time 200\
-   Also important to notice for later that those two lines cross\
-   Also printed here down below

```{r}
print(paste("Pr(T > 200) for prior 0:", signif(S_200_prior0, digits = 3)))
print(paste("Pr(T > 200) for prior 10:", signif(S_200_prior10, digits = 3)))
```

2.  Test whether there is a significant difference between the survival functions of the two `prior` groups and draw you conclusions. [**25 marks**]

#### Talking points, which rho?

-   First of all, as on the previous plot we've noticed that lines crossed, we concluded that we should use Wilcoxon test (rho=1) instead of Log-rank.\

```{r}
WT <- survdiff(Surv(time, status) ~ prior, data = veteran, rho = 1)
WT
```

#### Talking points diff

-   From this comparison, we got a p-value of 0.8 meaning that we failed to reject the null hypotesis as there is no statistically relevant difference between the survival functions for prior 0 (patient did not receive prior therapy) and prior 10 (patient received prior therapy)

Estimate the Weibull PH Model (using the **SurvRegCensCov** package) using the `age` and `prior` covariates. [**25 marks**]

#### Talking points Weibull PH Model

-   We gonna use the Veteran dataset that represents data from a randomised trial of two treatments for lung cancer.

-   Here we estimate the Weibull PH model using age and prior

-   We used Surv function to build the survival object

-   Put it explained by age and the factor variable prior

    \

```{r}
library(SurvRegCensCov)

WR <- WeibullReg(Surv(time, status) ~ age + factor(prior), data = veteran)
WR 
```

#### Talking points Weibull distribution

-   Gamma is \< 1 (0.867) so, hazard decreases over time.\
-   The model suggests that age has a negative effect on survival\
-   And having prior have a positive effect on survival\
-   But, age, with a p-value (0.317) and prior, with a p-value (0.349), are both above 0.05 so, we can conclude that they are not statistically relevant for the model

Create an R function estimating the hazard function of the Weibull PH model and, using the estimates obtained above, and plot the hazard curve for an individual with `age` = 30 and `prior` = 10. [**25 marks**]\

#### Talking points hazard function

-   Created a generic function to estimate hazard for this model\
-   Plotted hazard for an 30 years old individual with prior treatment\

```{r}

h_weib <- function(x, lambda, gamma, beta_age, beta_prior, x_age, x_prior) {
  lambda * gamma * x^(gamma-1) * exp(beta_age * x_age + beta_prior * x_prior)
}
lambdaVet <- WR$coef[1, 1]
gammaVet <- WR$coef[2, 1]
beta_ageVet <- WR$coef[3, 1]
beta_priorVet <- WR$coef[4, 1]


curve(
  h_weib(x, lambda = lambdaVet, gamma = gammaVet, beta_age = beta_ageVet, 
        beta_prior = beta_priorVet, x_age = 30, x_prior = 1), 
      from = 0, to = 1000, lwd = 3, 
      ylab = 'h(t)', xlab = 't', main = "Hazard function of the Weibull PH model for \nindividual with age 30 and prior treatment")
```

#### Talking points hazard chart

-   As expected, with gamma lower than 1, the hazard decreases over time
