---
html_document:
  mathjax: http://cdn.mathjax.org/mathjax/latest/MathJax.js?config=TeX-AMS-MML_HTMLorMML
author: "Mauricio Reis | D21125621"
title: "Assignment 2"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

Complete the following project using R Markdown and ensure that your code and output are readable in the html file.

### The `veteran` dataset
```{r}
library(survival)
?veteran
```

<br>

### Questions

<br>

1. Estimate and plot $S(t)$ for the two `prior` groups using the Kaplan-Meier estimator and find $\Pr(T > 200)$ for an individual with `prior` = 0 and `prior` = 10. [**25 marks**]
```{r}
SKM <- survfit(Surv(time, status) ~ prior, data = veteran)
summary(SKM)
```
```{r}
plot(SKM, 
     col = c("blue", "red"),
     lwd = 3,
     main = "S(t) for the two prior groups on veteran Dataset\n marking point of survival beyond t = 200")

legend("topright", 
       c("prior = 0", "prior = 10"), 
       lty = 1,
       lwd = 3,
       col = c("blue", "red"))
S_200_prior0 <- summary(SKM, times = 200)$surv[1]
S_200_prior10 <- summary(SKM, times = 200)$surv[2]
points(200, y = S_200_prior0, type = "p", col = "black", pch = 16)
points(200, y = S_200_prior10, type = "p", col = "black", pch = 16)

```

```{r}
print(paste("Pr(T > 200) for prior 0:", signif(S_200_prior0, digits = 3)))
print(paste("Pr(T > 200) for prior 10:", signif(S_200_prior10, digits = 3)))
```


2. Test whether there is a significant difference between the survival functions of the two `prior` groups and draw you conclusions. [**25 marks**]
```{r}
WT <- survdiff(Surv(time, status) ~ prior, data = veteran, rho = 1)
WT
```

*conclusions here.*  

First of all, as on the previous plot we've noticed that lines crossed, we concluded that we should use Wilcoxon test (rho=1) instead of Log-rank.  
From this comparison, we got a p-value of 0.8 meaning that we failed to reject the null hypotesis as there is no statistically relevant difference between the survival functions for prior 0 (patient did not receive prior therapy) and prior 10 (patient received prior therapy)  


3. Estimate the Weibull PH Model (using the **SurvRegCensCov** package) using the `age` and `prior` covariates. [**25 marks**]
```{r}
library(SurvRegCensCov)

WR <- WeibullReg(Surv(time, status) ~ age + factor(prior), data = veteran)
WR 
```

4. Create an R function estimating the hazard function of the Weibull PH model and,  using the estimates obtained above, and plot the hazard curve for an individual with `age` = 30 and `prior` = 10. [**25 marks**]
```{r}

h_weib <- function(x, lambda, gamma, beta_age, beta_prior, x_age, x_prior) {
  lambda * gamma * x^(gamma-1) * exp(beta_age * x_age + beta_prior * x_prior)
}
lambdaVet <- WR$coef[1, 1]
gammaVet <- WR$coef[2, 1]
beta_ageVet <- WR$coef[3, 1]
beta_priorVet <- WR$coef[4, 1]

curve(
  h_weib(x, lambda = lambdaVet, gamma = gammaVet, beta_age = beta_ageVet, 
        beta_prior = beta_priorVet, x_age = 30, x_prior = 1), 
      from = 0, to = 1000, lwd = 3, 
      ylab = 'h(t)', xlab = 't', main = "Hazard function of the Weibull PH model for \nindividual with age 30 and prior treatment")
```

