```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo=FALSE) 
```

```{r,echo=FALSE}
#set this file directory as current path
current_path = rstudioapi::getActiveDocumentContext()$path
setwd(dirname(current_path))
```

```{r}
#Install and load kable extra to render tables in a nicer way
#install.packages("kableExtra")
library(kableExtra)
#install and load tidyverse to do dataset manipulations
#install.packages("tidyverse")
library(tidyverse)
#read telecoms_churn csv file into a list in R and show basic 
telecoms_churn=read.csv("telecoms_churn.csv",header=T)

```

```{r, results='hide'}
# QUESTION 1 STARTS HERE
#1.a Fitting a model with all predictors except customerid as it's a unique identifier
attach(telecoms_churn)
fGender <- factor(gender)
fPartner <- factor(partner)
fPayMethod <- factor(paymentmethod)
fInternetservice <- factor(internetservice)
detach(telecoms_churn)
fit=glm(churn~monthlycharges+fGender+fPartner+fPayMethod+fInternetservice,
        family=binomial(),data=telecoms_churn)
summary(fit)
```

```{r, results='hide'}
# 1.b calculating probability of churn for a male, with partner, 
# monthly charges=70, internet service=fibre optic, payment method= credit card

nd=data.frame(monthlycharges=70,
              fGender="Male",
              fPartner="Yes",
              fPayMethod="Credit card (automatic)",
              fInternetservice="Fiber optic"
              )
#predict log odds
predLogOdds=predict(fit,type='link',newdata=nd, se.fit = TRUE)
paste("Log Odds:", predLogOdds$fit)
#predict probability
predProb=predict(fit,type='response',newdata=nd, se.fit = TRUE)
paste("Probability:", predProb$fit)
```

```{r, results='hide'}
#1.c.1 calculating the 95% Log Odds CI for the estimate above
uprLogOddsCI <- predLogOdds$fit + (1.96 * predLogOdds$se.fit)
lwrLogOddsCI <- predLogOdds$fit - (1.96 * predLogOdds$se.fit)
paste("Log Odds CI: [", lwrLogOddsCI,",",uprLogOddsCI,"]",sep = "")

```

```{r, results='hide'}
#1.c.1 calculating the 95% Probability CI for the estimate above
uprProbCI <- fit$family$linkinv(uprLogOddsCI)
lwrProbCI <- fit$family$linkinv(lwrLogOddsCI)
paste("Probability CI: [", lwrProbCI,",",uprProbCI,"]",sep = "")
```

# Telecoms Churn Data  

### Introduction  
```{r, results='hide'}
# just examining the data...
str(telecoms_churn)
prop.table(table(telecoms_churn$gender)) # table for genders 
prop.table(table(telecoms_churn[telecoms_churn$gender=="Male",]$partner)) #male w/ partners
prop.table(table(telecoms_churn[telecoms_churn$gender=="Female",]$partner)) 
# there is a higher proportion of female with partner, 
# and higher proportion of males without partner
count(telecoms_churn[!complete.cases(telecoms_churn),]) == 0 #there are no NAs
#how the charges differ between services?
mean(telecoms_churn$monthlycharges) ## mean 66.5
sd(telecoms_churn$monthlycharges) ## sd 29.9
mean(telecoms_churn$monthlycharges[telecoms_churn$internetservice == "No"]) # mean 21.3
sd(telecoms_churn$monthlycharges[telecoms_churn$internetservice == "No"]) # sd 2.22
mean(telecoms_churn$monthlycharges[telecoms_churn$internetservice == "Fiber optic"]) # mean 92.06
sd(telecoms_churn$monthlycharges[telecoms_churn$internetservice == "Fiber optic"]) # sd 12.22

```

The Telecoms Churn Data is consisted of 512 observations and 7 variables from which one of those, `churn`, represents whether the customer was lost within 2 years (value=1) or not (value=0). This dataset also has a field called monthlycharges that represents how much a customer is charged per month which is on average 66.5 but this sum varies according to the internet service being offered with "No" internet service monthlycharges mean being only 21.3.  
There is also a higher proportion of males 56%, than females. And, male customers without a partner are more common 53% whereas for female customers, it's more common to have a partner as of them 55% have.


### Model 1: Logistic Regression with all predictors (except customerid)
  
The model with all the predictors looks like this:
$$ \eta_i = \beta_0 + \beta_1 (monthlycharges_i) + \beta_2 \delta_{i,sexMale} + 
\beta_3 \delta_{i,partnerYes} + \beta_4 \delta_{i,payCreditCard} + \beta_5 \delta_{i,payEletronicCheck} \\
+ \beta_6 \delta_{i,payMailedCheck} + \beta_7 \delta_{i,serviceFiberOptic} + \beta_8 \delta_{i,serviceNo} $$  

With the following predictors seen on Table 1:
```{r}
summary(fit)$coefficients %>% 
  kbl(caption = "Table 1: Model 1 Coefficients") %>%
  kable_styling()
```


  - Using this model for a prediction, given an example of a male, living with his partner, monthly charges=70, internet service=fibre optic, payment method= credit card. The log odds of churning for him would be:

$$ \eta_i = 0.17470 - 0.04908 \times 70 + 0.21865 \times 1 - 0.67730 \times 1 + 0.72001 \times 1 + \\
1.72827 \times 0 + 0.94381 \times 0 + 1.76588 \times 1 - 2.78320 \times 0 \\
   \eta_i = -1.23366 $$
  
### Model 2 
Table 1 shows that, in Model 1, the p-value for gender male categorical predictor was very high: 0.410096. Let's do a custom hypothesis test to look into more detail if we really need this predictor in the model:  

1. $H_0: \beta_2 = 0$  
2. $H_\alpha:\beta_2\neq0$  
3. Custom hypothesis test result of customer with all same parameters except one is male and is another female:  

```{r}
# Comparing customer with all same parms except one is male and another female
library(multcomp)
L=cbind(0,0,-1,0,0,0,0,0,0)
glh=glht(fit,linfct=L)
summary(glh)
```
4. P-value is greater than 0.05 (0.41)  
5. We fail to reject the null hypothesis and conclude that gender is not significantly related to the response

- After the removal of gender, the new model looks like this now:  
$$ \eta_i = \beta_0 + \beta_1 (monthlycharges_i) + \beta_2 \delta_{i,partnerYes} + 
\beta_3 \delta_{i,payCreditCard} + \beta_4 \delta_{i,payEletronicCheck} + \\
\beta_5 \delta_{i,payMailedCheck} + \beta_6 \delta_{i,serviceFiberOptic} + \beta_7 \delta_{i,serviceNo}$$  
Also, Table 2 shows the coeficients for this new model

```{r}
fit2=update(fit,.~.-fGender)
summary(fit2)$coefficients %>% 
  kbl(caption = "Table 2: Model 2 Coefficients") %>%
  kable_styling()
```

```{r, results='hide'}
#check which is the minimum montly charge? It there someone not paying for the service?
min((telecoms_churn$monthlycharges))
#No
```

In this model, the intercept is estimating a customer without a partner, who pays with bank transfer and has DSL as his Internet Service and pays 0 monthly charge (that seems like very unlikely case in real life and doesn't have any occurrence on this dataset). Moving forward, this analisys will show how each of those predictors relate to the response.  

### Categorical Predictors levels and how they differ from each other

Let's first take a look into the payment method factor levels with high P-value, Credit card and Mailed check and how they compare with bank transfer. Let's see also if there is a different between them both and between Internet Service with Fiber Optic and with No Service.  

#### Customer with all same parameters except one uses Bank Transfer and the other Credit Card  
```{r}
getGlhPvalue <- function(aGlhResult) {
  return (paste("GLH p-value result:", summary(aGlhResult)$test$pvalues[1]))
}
# Comparing customer with all same parms except one uses Bank Transfer and the other Credit Card
L=cbind(0,0,0,-1,0,0,0,0)
glh=glht(fit2,linfct=L)
getGlhPvalue(glh)

```
As shown in the output of "Simultaneous Tests for General Linear Hypotheses
", with P-value higher than 0.05 (0.221), it's possible to conclude that Bank Transfer and Credit Card predictors difference isn't statistically relevant.
  

#### Customer with all same parameters except one uses Bank Transfer and the other Mailed Check  

```{r}
# Comparing customer with all same parms except one uses Bank Transfer and the other Mailed Check
L=cbind(0,0,0,0,0,-1,0,0)
glh=glht(fit2,linfct=L)
getGlhPvalue(glh)

```
With P-value higher than 0.05 (0.105), it's possible to conclude that Bank Transfer and Mailed Check predictors difference also isn't statistically relevant.

#### Customer with all same parameters except one uses Credit Card and the other Mailed Check  
```{r}

# Comparing customer with all same parms except one uses Credit Card and the other Mailed Check
L=cbind(0,0,0,1,0,-1,0,0)
glh=glht(fit2,linfct=L)
getGlhPvalue(glh)
```
With P-value higher than 0.05 (0.624), it's possible to conclude that Credit Card and Mailed Check predictors difference isn't statistically relevant.  

#### Customer with all same parameters except one has Fiber Optic service and other has No Internet Service
```{r}

# Comparing customer with all same parms except one hasFiber Optic service and other 
#has No Internet Service
L=cbind(0,0,0,0,0,0,1,-1)
glh=glht(fit2,linfct=L)
getGlhPvalue(glh)

```
With a p-value way below 0.05 (7.577e-07), it's possible to conclude that the difference between Fiber Optic and No service is statistically relevant.


### Odds Ratios

#### Odds Ratio for a customer without partner over a customer with a partner:
$$logOddsRatio= -\beta_2 = -(-0.69949) = 0.69949$$
$$oddsRatio= e^{0.69949}=2.012726$$
As shown above, the odds for a customer without a partner to churn are higher than in the case of a customer without a partner.  

#### Odds Ratio for a customer using bank transfer over a customer using electronic check
$$logOddsRatio= -\beta_4 = -1.71683$$
$$ oddsRatio= e^{-1.71683}=0.1796347 $$ 
As shown above, the odds of churn for a customer using bank transfer are lower than for a customer using eletronic check.

#### Odds Ratio for a customer that has DSL over a customer that  has Fiber optic service.
$$ logOddsRatio= -\beta_6 = -1.74224 $$ 
$$ oddsRatio= e^{-1.74224}=0.1751277 $$ 
The odds of churn for a customer using DSL are smaller than for a customer using Fiber optic.

#### Odds Ratio for a customer that has DSL over a customer that  has No Internet service
$$ logOddsRatio= -\beta_7 = -(-2.78252) = 2.78252 $$ 
$$ oddsRatio= e^{2.78252}=16.15969$$ 
The odds of churn for a customer with DSL are dramatically higher than for a customer with No Internet service.

#### Odds Ratio for a customer that has Fiber optic over a customer that  has No Internet service
$$ logOddsRatio= \beta6-\beta_7 = 1.74224 -(-2.78252) = 4.52476 $$ 
$$ oddsRatio= e^{4.52476}=92.27378$$
The odds of churn for a customer with Fiber optic are dramatically higher than for a customer with No Internet service.

### Conclusions and limitations of the model 2

As risk factors for churn we can consider that:  

- The higher the Monthly charges, the less likely the customers are to churn, which might be true but, may suggest a need to a interaction term to be added to this model or even that we it's missing some important piece of information from customers that might affect montlycharges like time since contract has started. Other possibility, depending the business needs, is that this dataset could be divided and analysed separately for each type of service since their average monthly charges vary considerably and their customer profile and competitiviness could also vary. 

- Customers with a partner are less likely to churn also, according to this model is safe to make the assumption that not having a partner is a risk factor to leave the service. 

- Customers who pay with Eletronic check are less likely to churn than the customers with other forms of payments. This analysis didn't find any evidence that the other remaining types payments are statiscally different from each other, all can be considered risk factors.

- Customers with Fiber optic service are more likely to churn than with the other services and customers without Internet service are considerably less likely to churn. This analysis has shown that having any internet service, and specially Fiber optic service, is one of risk factors to churn.  



\newpage  

## R Code:  

```{r, echo=TRUE, eval = FALSE}
#set this file directory as current path
current_path = rstudioapi::getActiveDocumentContext()$path
setwd(dirname(current_path))
```

```{r, echo=TRUE, eval = FALSE}
#Install and load kable extra to render tables in a nicer way
#install.packages("kableExtra")
library(kableExtra)
#install and load tidyverse to do dataset manipulations
#install.packages("tidyverse")
library(tidyverse)
#read telecoms_churn csv file into a list in R and show basic 
telecoms_churn=read.csv("telecoms_churn.csv",header=T)

```

```{r, echo=TRUE, eval = FALSE}
# QUESTION 1 STARTS HERE
#1.a Fitting a model with all predictors except customerid as it's a unique identifier
attach(telecoms_churn)
fGender <- factor(gender)
fPartner <- factor(partner)
fPayMethod <- factor(paymentmethod)
fInternetservice <- factor(internetservice)
detach(telecoms_churn)
fit=glm(churn~monthlycharges+fGender+fPartner+fPayMethod+fInternetservice,
        family=binomial(),data=telecoms_churn)
summary(fit)
```

```{r, echo=TRUE, eval = FALSE}
# 1.b calculating probability of churn for a male, with partner, 
# monthly charges=70, internet service=fibre optic, payment method= credit card

nd=data.frame(monthlycharges=70,
              fGender="Male",
              fPartner="Yes",
              fPayMethod="Credit card (automatic)",
              fInternetservice="Fiber optic"
              )
#predict log odds
predLogOdds=predict(fit,type='link',newdata=nd, se.fit = TRUE)
paste("Log Odds:", predLogOdds$fit)
#predict probability
predProb=predict(fit,type='response',newdata=nd, se.fit = TRUE)
paste("Probability:", predProb$fit)
```

```{r, echo=TRUE, eval = FALSE}
#1.c.1 calculating the 95% Log Odds CI for the estimate above
uprLogOddsCI <- predLogOdds$fit + (1.96 * predLogOdds$se.fit)
lwrLogOddsCI <- predLogOdds$fit - (1.96 * predLogOdds$se.fit)
paste("Log Odds CI: [", lwrLogOddsCI,",",uprLogOddsCI,"]",sep = "")

```

```{r, echo=TRUE, eval = FALSE}
#1.c.1 calculating the 95% Probability CI for the estimate above
uprProbCI <- fit$family$linkinv(uprLogOddsCI)
lwrProbCI <- fit$family$linkinv(lwrLogOddsCI)
paste("Probability CI: [", lwrProbCI,",",uprProbCI,"]",sep = "")
```

```{r, echo=TRUE, eval = FALSE}
# just examining the data...
str(telecoms_churn)
prop.table(table(telecoms_churn$gender)) # table for genders 
prop.table(table(telecoms_churn[telecoms_churn$gender=="Male",]$partner)) #male w/ partners
prop.table(table(telecoms_churn[telecoms_churn$gender=="Female",]$partner)) 
# there is a higher proportion of female with partner, 
# and higher proportion of males without partner
count(telecoms_churn[!complete.cases(telecoms_churn),]) == 0 #there are no NAs
#how the charges differ between services?
mean(telecoms_churn$monthlycharges) ## mean 66.5
sd(telecoms_churn$monthlycharges) ## sd 29.9
mean(telecoms_churn$monthlycharges[telecoms_churn$internetservice == "No"]) # mean 21.3
sd(telecoms_churn$monthlycharges[telecoms_churn$internetservice == "No"]) # sd 2.22
mean(telecoms_churn$monthlycharges[telecoms_churn$internetservice == "Fiber optic"]) # mean 92.06
sd(telecoms_churn$monthlycharges[telecoms_churn$internetservice == "Fiber optic"]) # sd 12.22
```

```{r, echo=TRUE, eval = FALSE}
summary(fit)$coefficients %>% 
  kbl(caption = "Table 1: Model 1 Coefficients") %>%
  kable_styling()
```


```{r, echo=TRUE, eval = FALSE}
# Comparing customer with all same parms except one is male and another female
library(multcomp)
L=cbind(0,0,-1,0,0,0,0,0,0)
glh=glht(fit,linfct=L)
summary(glh)
```

```{r, echo=TRUE, eval = FALSE}
fit2=update(fit,.~.-fGender)
summary(fit2)$coefficients %>% 
  kbl(caption = "Table 2: Model 2 Coefficients") %>%
  kable_styling()
```

```{r, echo=TRUE, eval = FALSE}
#check which is the minimum montly charge? It there someone not paying for the service?
min((telecoms_churn$monthlycharges))
#No
```

```{r, echo=TRUE, eval = FALSE}
getGlhPvalue <- function(aGlhResult) {
  return (paste("GLH p-value result:", summary(aGlhResult)$test$pvalues[1]))
}
# Comparing customer with all same parms except one uses Bank Transfer and the other Credit Card
L=cbind(0,0,0,-1,0,0,0,0)
glh=glht(fit2,linfct=L)
getGlhPvalue(glh)
```

```{r, echo=TRUE, eval = FALSE}
# Comparing customer with all same parms except one uses Bank Transfer and the other Mailed Check
L=cbind(0,0,0,0,0,-1,0,0)
glh=glht(fit2,linfct=L)
getGlhPvalue(glh)
```

```{r, echo=TRUE, eval = FALSE}
# Comparing customer with all same parms except one uses Credit Card and the other Mailed Check
L=cbind(0,0,0,1,0,-1,0,0)
glh=glht(fit2,linfct=L)
getGlhPvalue(glh)
```

```{r, echo=TRUE, eval = FALSE}
# Comparing customer with all same parms except one hasFiber Optic service and other 
#has No Internet Service
L=cbind(0,0,0,0,0,0,1,-1)
glh=glht(fit2,linfct=L)
getGlhPvalue(glh)
```

