# Regression 8A Code

```{r}
current_path = rstudioapi::getActiveDocumentContext()$path
setwd(dirname(current_path))
```

```{r}
titanic=read.csv('titanic.csv',header=T)
titanic=titanic[complete.cases(titanic),]
colnames(titanic)=tolower(colnames(titanic))
fit=glm(survived~age+factor(sex)+factor(pclass),family=binomial(),data=titanic)
summary(fit)
```
## customised hypothesis tests
### compare second class with 3rd class

```{r}
library(multcomp)
L=cbind(0,0,0,1,-1)
glh=glht(fit,linfct=L)
summary(glh) # test
```
```{r}
confint(glh)$confint # CI on linear predictor scale (log odds)

```

```{r}
exp(confint(glh)$confint) # CI on OR scale 
```
## compare a woman in 3rd class with a man in 2nd class?

```{r}
L=cbind(0,0,-1,-1,1)
glh2=glht(fit,linfct=L)
summary(glh2) # test

```

```{r}
exp(confint(glh)$confint) # CI on OR scale 

```
## compare a woman in 3rd class with a man in 1st class?

```{r}
L=cbind(0,0,-1,0,1)
glh3=glht(fit,linfct=L)
summary(glh3) # test
```

```{r}
exp(confint(glh3)$confint) # CI on OR scale 

```

# Likelihood Ratio Tests (LRTs) for Logistic regression
```{r}
neuralgia=read.table("neuralgia.txt",header=T)
colnames(neuralgia)=tolower(colnames(neuralgia))
head(neuralgia)
```
### Fit the model and get the log likelihood

```{r}
fit1=glm(pain~age+duration+factor(sex)+factor(treatment),family=binomial(),data=neuralgia)
summary(fit1)

```

```{r}
logLik(fit1)
```

### Testing the global null hypothesis
```{r}
fit0=update(fit1,.~1)
logLik(fit0)
```

```{r}
C=-2*(logLik(fit0)-logLik(fit1))
C
```

```{r}
pvalue=1-pchisq(C,5)
pvalue
```
### Quicker method…
```{r}
anova(fit0,fit1,test='LRT')

```
## Testing for the need for sex and duration in the model

```{r}
anova(update(fit1,.~.-duration-factor(sex)),fit1,test='LRT')

```

### model building via LRT
```{r}
drop1(fit1,test='LRT')

```

```{r}
fit2=update(fit1,.~.-duration)
drop1(fit2,test='LRT')
```

```{r}
summary(fit2)

```

