---
html_document:
  mathjax: http://cdn.mathjax.org/mathjax/latest/MathJax.js?config=TeX-AMS-MML_HTMLorMML
author: "Mauricio Reis | D21125621"
title: "Assignment 1"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, cache = TRUE)
```

Complete the following project using R Markdown and ensure that your code and output are readable in the html file.

### The `gangs` network dataset

Criminal (sub)network of 67 gang members of Balkan crime groups in Italy. The following vertex attributes are provided:

* `nationality`: nationality of the criminal (1 = Serbian, 2 = Romanian, 3 = Greek, 4 = Bulgarian);

* `sex`: sex of the criminal (1 = female, 2 = male);

* `role`: role of the criminal (1 = leader, 2 = 'thug', 3 = 'lackie').

```{r message=FALSE}
library(statnet)
load(url("https://acaimo.github.io/Bergm/network_data/gangs.Rdata"))
```

```{r}
#helper functions 
nationalityList <- c('Serbian', 'Romanian', 'Greek', 'Bulgarian')
nationalityColList <- c('Red', 'Yellow', 'Blue', 'Green')
sexList <- c('female', 'male')
sexColList <- c('pink', 'blue')
roleList <- c('leader', 'thug', 'lackie')
roleColList <- c('gold', 'grey', 'brown')
describeNode <- function(nodeNumber) {

  nationality <- (y %v% "nationality")[nodeNumber]
  sex <- (y %v% "sex")[nodeNumber]
  role <- (y %v% "role")[nodeNumber]
  paste("Node ", nodeNumber, " - ", 
        "nationality: ", nationalityList[nationality], ", ",
        "sex: ",  sexList[sex], ", ",
        "role: ", roleList[role], sep = "")
}
```


### Description

Calculate and interpret:

1. the density of the network;  [**10 marks**]
```{r}
y <- gangs
summary(y ~ density)

```

*Comment*  
This value represents the number of edges divided by the number of dyads on the network and serves to assert if a network is dense or sparse. This network is sparse with $density < 0.5$  
  
---  

2. in- and out- degree distribution, plot the respective histograms, and identify the node(s) with the highest degree measures;  [**10 marks**]
```{r}
id_i <- degree(y, cmode = "indegree")
od_i <- degree(y, cmode = "outdegree")
numVertex <- length(network.vertex.names(y))
par(mfrow = c(1, 2))


barplot(table(id_i) / numVertex, 
        main = "In-degree distribution",
        col = "skyblue",
        xlab = "In-degree",
        ylab = "Relative frequency")

barplot(table(od_i) / numVertex, 
        main = "Out-degree distribution",
        col = "skyblue",
        xlab = "Out-degree",
        ylab = "Relative frequency")

nodesWithMaximumInDegree <- which(id_i == max(id_i))
nodesWithMaximumOutDegree <- which(od_i == max(od_i))
print("The node(s) with maximum In Degree is/are:")
print(sapply(nodesWithMaximumInDegree,describeNode))
print("The node(s) with maximum Out Degree is/are:")
print(sapply(nodesWithMaximumOutDegree,describeNode))
```

*Comment*  

From the histogram above we can notice that the most common by far for a node in the network is to receive (in) 0 connections. In the case of a node pointing out to another node (out), the most common is point out to 1 other node closely followed by 0.  

We also discovered that the nodes with the maximum In Degree are 30 (a Serbian female lackie) and 57 (a Bulgarian female thug) and the node with the maxium Out degree is also the 57.

---  

3. the geodesic distances of the network; [**10 marks**]
```{r fig.height=7, fig.width=7}
gd <- geodist(y)$gdist

cp <- hcl.colors(nlevels(factor(gd)), "Teal")

heatmap(gd,
        Rowv = NA, Colv = "Rowv", 
        symm = TRUE, 
        col = cp, 
        frame = TRUE)
```

*Comment*  

As you can see in the heapmap above, there is not one single node in this network that connect with all the other nodes (we at least a few blank spaces on every line). We can also notice the diagonal dark line that shows for each position where the node references itself. The varied colours show that we have some nodes connecting with others through a long path (lighter colours) and others through a short path (darker colours)  

  
---  

4. the betweenness centrality (and plot it) and identify the nodes with the highest values of this measure; [**10 marks**]
```{r fig.height=7, fig.width=8}
set.seed(2)
par(mar = rep(0, 4))
gplot.target(y[, ], 
             betweenness(y), 
             gmode = "graph", 
             label = 1:length(network.vertex.names(y)), 
             vertex.cex = 0.3,
             edge.col = 'lightgrey',
             label.cex = 1,
             circ.lab.cex = 0.5)

```
```{r}
describeNode(which.max(betweenness(y)))
```

*Comment*  
The node with the highest value for betweenness centrality is node 30 (a Serbian female lackie), followed by node 57 (a Bulgarian female thug). This means that those two gang members are the top 2 the shortest paths to connect other gang members (in terms of how many can go through this path, not necessarily that everyone can or has this as the shortest path to any other member).
  
  
---  

5. the clustering coefficient of the network; [**10 marks**]
```{r}
gtrans(y[, ],           
       mode = "digraph")
```

*Comment*  
Clustering coeficient is the number of triangles in a network divided by the number of connected triples. 

Given that a clustering coeficient of 1 means that on our context "if gang member x is connected with member y, and member y is connect with member z, so member x is connected with member z", this value 0.0429 means that this is not that case in most of the times

  
---  

6. the **edgewise** reciprocity between criminals with the same nationality (hint: `mutual(same  = ...)`). [**10 marks**]
```{r}
summary(y ~ edges + mutual(same = "nationality") + nodematch("nationality"))
reciprocity <- 14/73
reciprocityNationalityOnly <- 14/59
print(paste("Reciprocity", reciprocity))
print(paste("Reciprocity considering nationality only", reciprocityNationalityOnly))
```

*Comment*  

Reciprocity is the number of mutual edges divided by the number of edges in the network. For the gangs dataset we found 14  mutual edges among a total of 73 edges in the network giving a reciprocity of 0.192. If we consider only the edges that have a matching nationality instead of all edges though, then this number goes up to 0.237.
  
---  

### Visualisation 

7. Plot two network graphs by setting the vertex sizes proportional on their in- and out-degree respectively using the "fruchtermanreingold" layout (include vertex labels); [**10 marks**]
```{r fig.height=7, fig.width=8}
set.seed(2)
par(mar = rep(1, 4)) 
#id_i <- degree(y, cmode = "indegree")
#od_i <- degree(y, cmode = "outdegree")

plot(y, 
     boxed.labels = TRUE, # nodal labels within boxes  
     label = id_i,    # representing in degree
     mode = "fruchtermanreingold", 
     edge.col = "grey",
     vertex.cex = id_i /2,
     main = "Vertex Sizes proportional to In Degree")
plot(y, 
     boxed.labels = TRUE, # nodal labels within boxes  
     label = od_i,    # representing out degree
     mode = "fruchtermanreingold", 
     edge.col = "grey",
     vertex.cex = id_i /2,
     main = "Vertex Sizes proportional to Out Degree")
```

*Comment*  
The charts above shows that many nodes don't have neither in neither out connections and, looking at the in degree chart, we can see now with more clarity the two nodes with biggest in degree (7) proportional to the in degree in of the others, in the out degree chart we can see the proportion of the biggest out degree node (9) too but. we noticed here that there is second node with very high out degree compared to the others (8), which we didn't noticed before without a plot.
  
  
---  
      
8. Plot three network graphs (with legend) setting a different the vertex colour for each level of the nodal attributes using the "kamadakawai" layout. [**10 marks**]
```{r fig.height=7, fig.width=7}
set.seed(5)
par(mar = rep(1, 4)) 
plot(y, 
     label = 1:numVertex,
     label.cex = 0.5,
     vertex.cex = 1,
     edge.col = "grey",
     vertex.col = nationalityColList[y %v% "nationality"],
     main = "Gangs network with nationality described by color",
     mode = "kamadakawai")

legend("topright", 
       pt.bg  = nationalityColList[sort(unique(y %v% "nationality"))], 
       pt.cex = 1.2,
       pch    = 21, 
       legend = nationalityList[sort(unique(y %v% "nationality"))], 
       title  = 'Nationality')
```
```{r}
set.seed(5)
par(mar = rep(1, 4)) 
plot(y, 
     label = 1:numVertex,
     label.cex = 0.5,
     vertex.cex = 1,
     edge.col = "grey",
     vertex.col = sexColList[y %v% "sex"],
     main = "Gangs network with sex described by color",
     mode = "kamadakawai")

legend("topright", 
       pt.bg  = sexColList[sort(unique(y %v% "sex"))], 
       pt.cex = 1.2,
       pch    = 21, 
       legend = sexList[sort(unique(y %v% "sex"))], 
       title  = 'sex')

```
```{r}
set.seed(4)
par(mar = rep(1, 4)) 
plot(y, 
     label = 1:numVertex,
     label.cex = 0.5,
     vertex.cex = 1,
     edge.col = "grey",
     vertex.col = roleColList[y %v% "role"],
     main = "Gangs network with role described by color",
     mode = "kamadakawai")

legend("topright", 
       pt.bg  = roleColList[sort(unique(y %v% "role"))], 
       pt.cex = 1.2,
       pch    = 21, 
       legend = roleList[sort(unique(y %v% "role"))], 
       title  = 'role')

```


*Comment*  
On nationality plot, apart from some isolated individuals, we can see how members of a same nationality are clustered together, with the exception of the greeks, who seem to have 3 separate small groups. Also Romanians only seem to have direct links with Serbians.  

On the sex plot, we actually don't see clustering by sex as the few females in the plot do not have direct links between each other.  
  
On the role plot might suggest a hyerarchy of some sorts, we see that lakies are the majority. Thugs are usually surrounded by lackies, and there is only one leader that has two in edges from thugs but, for some reason, has also a bidiretional edge with a lackie.
  

### Statistical modelling  [**20 marks**]

Estimate, interpret and assess the exponential-family random graph model specified by the following network effects:

  + density;

  + reciprocity;
  
  + reciprocity between actors with the same role;
  
  + homophily between actors with the same nationality.

```{r}
#ergmodel_1 <- y ~ 
#  edges + # density
#  mutual + # reciprocity 
#  mutual(same = "role") + # reciprocity between actors with the same role
#  nodematch('nationality') # homophily between actors with the same nationality


ergmodel_1 <- y ~ edges + mutual + mutual(same = "role") + nodematch('nationality')
#TODO: this is failing intermettently, find a solution to fix it.
mle.ergmodel_1 <- ergm(ergmodel_1)

summary(mle.ergmodel_1)
```

*Comment*  
```{r eval = FALSE}
#ATTENTION! the model above is failing intermittently 
#and not working at all when using knit
#this is the actual model summary when it works:
Call:
ergm(formula = ergmodel_1)

Monte Carlo Maximum Likelihood Results:

                      Estimate Std. Error MCMC % z value Pr(>|z|)    
edges                  -5.6006     0.2585      0 -21.666  < 1e-04 ***
mutual                  4.6172     0.4440      0  10.399  < 1e-04 ***
mutual.role            -2.4866     0.7612      0  -3.266  0.00109 ** 
nodematch.nationality   1.9692     0.2931      0   6.718  < 1e-04 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

     Null Deviance: 6130.2  on 4422  degrees of freedom
 Residual Deviance:  556.2  on 4418  degrees of freedom
 
AIC: 564.2  BIC: 589.8  (Smaller is better. MC Std. Err. = 0.6127)
```

From a first look we can say that, according to this model, this network seems to be sparse with edges predictor -5.6216 (way below 0.5), and very a significant P-value 1e-04. In fact, all predictors in this model are very significant.  
Another interesting fact is that the mutual reciprocity effect is positive but, for same role, the mutual reciprocity effect is negative. In other words, is more likely for gang member A to have a link back from gang member B if gang member A has a link to gang member B except when they are in the same role, in this case, it's less likely.  
Having the same nationality also seems to positively influence the likelihood of have a link between two gang members  


Goodness of fit:  
```{r}
gof(mle.ergmodel_1)
```
```{r eval=FALSE}
#ATTENTION! as the model above is failing on knit, this is the actual 
#goodness of fit fot he model when it works 
#(notice that here mutual role actually has a p-value of 1 instead of 0)
Goodness-of-fit for in-degree 

         obs min  mean max MC p-value
idegree0  30  14 21.94  36       0.14
idegree1  17  14 24.88  35       0.06
idegree2  12   6 13.02  22       0.96
idegree3   6   0  5.15  11       0.82
idegree4   0   0  1.62   6       0.38
idegree5   0   0  0.32   2       1.00
idegree6   0   0  0.07   1       1.00
idegree7   2   0  0.00   0       0.00

Goodness-of-fit for out-degree 

         obs min  mean max MC p-value
odegree0  26  12 22.77  39       0.52
odegree1  27  14 23.39  34       0.42
odegree2   7   6 13.59  21       0.08
odegree3   5   0  5.22  12       1.00
odegree4   0   0  1.60   5       0.50
odegree5   0   0  0.35   3       1.00
odegree6   0   0  0.07   2       1.00
odegree7   0   0  0.01   1       1.00
odegree8   1   0  0.00   0       0.00
odegree9   1   0  0.00   0       0.00

Goodness-of-fit for edgewise shared partner 

     obs min  mean max MC p-value
esp0  66  38 71.41  90       0.52
esp1   7   0  3.35  13       0.38
esp2   0   0  0.11   2       1.00

Goodness-of-fit for minimum geodesic distance 

     obs  min    mean  max MC p-value
1     73   38   74.87   98       0.82
2    154   17   80.33  143       0.00
3    127    5   76.38  164       0.18
4    101    2   65.96  180       0.42
5     86    0   53.54  194       0.44
6     54    0   41.05  203       0.58
7     37    0   29.17  197       0.62
8     16    0   20.09  163       0.80
9      7    0   13.31  130       0.92
10     3    0    9.00  105       0.88
11     0    0    5.36   67       1.00
12     0    0    2.83   41       1.00
13     0    0    1.56   33       1.00
14     0    0    0.66   25       1.00
15     0    0    0.29   18       1.00
16     0    0    0.13   12       1.00
17     0    0    0.08    8       1.00
18     0    0    0.05    5       1.00
19     0    0    0.02    2       1.00
20     0    0    0.01    1       1.00
Inf 3764 2783 3947.31 4360       0.48

Goodness-of-fit for model statistics 

                      obs min  mean max MC p-value
edges                  73  38 74.87  98       0.82
mutual                 17   8 17.64  27       0.92
mutual.role             2   0  1.94   6       1.00
nodematch.nationality  59  31 60.40  78       0.92
```

In a simulated network using our model, for in degree, we noticed that we just don't have relevant results for idegree1 and idegree7 with p-value < 0.05. And for out degrees, we would not have relevant results fot degree 8 and 9.  
For edgewise shared partners our model yelds relevant results for 0, 1 and 2.  
And for geodistic distance we just didn't have relevant results for 2. 
It's seems like, overall, the predictors chosen for this network yelded a fairly good model.




  





