---
html_document:
  mathjax: http://cdn.mathjax.org/mathjax/latest/MathJax.js?config=TeX-AMS-MML_HTMLorMML
output: html_notebook
title: "Assignment 1"
author: "Mauricio Reis | 0000000"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, cache = TRUE)
```

<br>

### Two dice $D_1$ and $D_2$ are rolled and let $S$ be the sum $S = D_1 + D_2.$
```{r}
# Exercise setup:
die1 <- c(1:6)
die2 <- c(1:6)
rolls <- expand.grid(die1, die2)
R <- data.frame(rolls, rowSums(rolls))
colnames(R) <- c("D_1", "D_2", "S")
totalDifferentRolls <- dim(R)[1]
```
#### 1. Find $\Pr(D_1 < 4\ \cup\ S > 9)$ [**10 marks**]
```{r}
sum(R$D_1 < 4 | R$S > 9) / totalDifferentRolls
```

#### 2. Find $\Pr(D_1 < 4\ |\ S = 8)$ [**10 marks**]
```{r}
sum(R$D_1 < 4 & R$S == 8) / sum(R$S == 8)
```

### Suppose $X \sim Binomial(n = 10, p = 0.3).$

#### 3. Find $\Pr(X \le 6 \cap 2 < X < 8).$ [**10 marks**]
```{r}
# insert code
xSmallerOrEqualToSix <- c(1:6)
xBetweenTwoAndEight<- c(3:7)
intersectionEx3 <- intersect(xSmallerOrEqualToSix, xBetweenTwoAndEight)
ex3Answer <- 0
for (xValue in intersectionEx3) {
  ex3Answer <- ex3Answer + dbinom(xValue, size = 10, prob = 0.3)
}
print(ex3Answer)

#TODO: Or should I do?
pbinom(6, size = 10, prob = 0.3) * (pbinom(7, size = 10, prob = 0.3) - pbinom(2,size = 10, prob = 0.3))
```

#### 4. Find $\Pr(X < 4\ |\ X \le 6).$ [**10 marks**]
```{r}
xSmallerThanFour <- c(1:3)
xSmallerOrEqualToSix <- c(1:6)
intersectionEx4 <- intersect(xSmallerThanFour, xSmallerOrEqualToSix)
intersectionProb <- 0
for (xValue in intersectionEx4) {
  intersectionProb <- intersectionProb + dbinom(xValue, size = 10, prob = 0.3)
}
xSmallerOrEqualToSixProb <- pbinom(6, size = 10, prob = 0.3)
intersectionProb / xSmallerOrEqualToSixProb
```

### Suppose $X \sim Poisson(\lambda = 5).$

#### 5. Find $\Pr(3 < X \le 7).$ [**10 marks**]
```{r}
# insert code
```

#### 6. Find $\Pr(X < 3 \cup X \ge 7).$ [**10 marks**]
```{r}
# insert code
```

### Create a function [**40 marks**]

Suppose $X \sim Poisson(\lambda).$ Create a function which plots the pmf and cdf for the interval $[x_1 - 5, x_1 + 5]$ and calculate the $p(x_1)$ and $F(x_1)$ for a given value of $x_1.$

* Input: $x_1$ and $\lambda;$
* Output: $p(x_1)$ and $F(x_1)$
* Extra feature: stop the function if $\lambda$ is negative and print a warning message. 
[**40 marks**]
```{r}
# insert code
```



