---
title: "Assignment 1"
author: "Mauricio de Oliveira Reis | D21125621"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, cache = TRUE)
```

<br>

### Two dice $D_1$ and $D_2$ are rolled and let $S$ be the sum $S = D_1 + D_2.$
  
#### 1. Find $\Pr(D_1 < 4\ \cup\ S > 9)$ [**10 marks**]  
```{r}
# Exercise 1 and 2 data setup:
die1 <- c(1:6)
die2 <- c(1:6)
rolls <- expand.grid(die1, die2) 
R <- data.frame(rolls, rowSums(rolls))
colnames(R) <- c("D_1", "D_2", "S")
totalDifferentRolls <- dim(R)[1]
#Exercise 1:
answer1 <- sum(R$D_1 < 4 | R$S > 9) / totalDifferentRolls
answer1
```

#### 2. Find $\Pr(D_1 < 4\ |\ S = 8)$ [**10 marks**]  

```{r}
answer2 <- sum(R$D_1 < 4 & R$S == 8) / sum(R$S == 8)
answer2
```
  
  

### Suppose $X \sim Binomial(n = 10, p = 0.3).$  
#### 3. Find $\Pr(X \le 6 \cap 2 < X < 8).$ [**10 marks**]  
```{r}
#data setup exercise 3 and 4
n <- 10
p <- 0.3
#exercise 3
xSmallerEqualSix <- c(0:6)
xBetweenTwoAndEight <- c(3:7)
intersection <- intersect(xSmallerEqualSix, xBetweenTwoAndEight)
answer3 <- sum(dbinom(intersection,n,p))
answer3
```


#### 4. Find $\Pr(X < 4\ |\ X \le 6).$ [**10 marks**]  
$$\Pr(X < 4\ \cap X \le 6) -> Pr(X < 4)$$
```{r}
#Pr(X<4 And X<= 6) / Pr(X<=6) -> Pr(X<4) / Pr(X<=6) 
probSmallerThanFour <- pbinom(3, n, p)
probSmallerEqualSix <- pbinom(6, n, p)
answer4 <- probSmallerThanFour / probSmallerEqualSix
answer4
```


### Suppose $X \sim Poisson(\lambda = 5).$  
#### 5. Find $\Pr(3 < X \le 7).$ [**10 marks**]  
```{r}
#data setup exercise 5 and 6
lambda <- 5

# exercise 5
xBetweenThreeAndSevenInclusive <- c(4:7)
answer5 <- sum(dpois(xBetweenThreeAndSevenInclusive, lambda))
answer5
```

#### 6. Find $\Pr(X < 3 \cup X \ge 7).$ [**10 marks**]  
$$ Pr(X < 3 \cap X \ge 7) = 0$$
```{r}
# insert code
probSmallerThanThree <- ppois(2,lambda)
probBiggerOrEqualSeven <- ppois(6, lambda, lower.tail = FALSE)
probOfIntersection <- 0
answer6 <- probSmallerThanThree + probBiggerOrEqualSeven - probOfIntersection
answer6
```



### Create a function [**40 marks**]

Suppose $X \sim Poisson(\lambda).$ Create a function which plots the pmf and cdf for the interval $[x_1 - 5, x_1 + 5]$ and calculate the $p(x_1)$ and $F(x_1)$ for a given value of $x_1.$

* Input: $x_1$ and $\lambda;$
* Output: $p(x_1)$ and $F(x_1)$
* Extra feature: stop the function if $\lambda$ is negative and print a warning message.   
[**40 marks**]
```{r}
poisson <- function(x1, lambda, plot = FALSE) {
  if (lambda <= 0) stop('lambda must be a positive number!')
  p_x <- dpois(x1, lambda)
  F_x <- ppois(x1, lambda)
  if (plot == TRUE) {
    par(mfrow = c(1, 2))
    INTERVAL_SIZE <- 5
    L <- x1 - INTERVAL_SIZE
    U <- x1 + INTERVAL_SIZE
    x <- c(L : U)
    pmf <- dpois(x, lambda)
    plot(x, pmf, lwd = 3, type = 'h', main = 'pmf of X', ylab = 'p(x)')
    
    cdf <- stepfun(x, c(ppois(L - 1, lambda), ppois(x, lambda)))
    plot.stepfun(cdf, 
                 verticals = FALSE, 
                 do.points = TRUE, 
                 pch = 16, lwd = 3,
                 main = 'cdf of X',
                 ylab = 'F(x)')
  }  
  list(p_x = p_x,
       F_x = F_x)
}
#poisson(10, 0) #This should throw an error
poisson(10, 5) #Only returns p_x and F_x
poisson(10, 5, TRUE) #Prints plot and returns p_x and F_x
```



