---
html_document:
  mathjax: http://cdn.mathjax.org/mathjax/latest/MathJax.js?config=TeX-AMS-MML_HTMLorMML
author: "Mauricio de Oliveira Reis | D21125621"
title: "Assignment 2"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, cache = TRUE)
```

<br>

### Suppose $X$ is a continuous r.v. with pdf: $f(x) = 3x^2, \ -1 < x < 0.$

#### 1. Plot the pdf of $X$ and verify that $f(x)$ is a pdf. [**10 marks**]
```{r}
f_x <- function(x) 3 * (x^2)
curve(f_x, from = -1, to = 0, ylab = 'f(x)', 
      main = 'pdf of X', lwd = 3)
f_all <- integrate(f_x, lower = -1, upper = 0)
if(f_all$value == 1) {
  print("Question 1 answer is: this is a pdf")
} else print("Question 1 answer is: this is NOT a pdf")
```

#### 2. Find $\Pr(X < -0.6\ |\ X > -0.8)$ [**10 marks**]
$$ X < -0.6\ \cap  X > -0.8 = -0.8 < X < -0.6  $$ 
```{r}
F_Between0.8And0.6 <- integrate(f_x, lower = -0.8, upper = -0.6)$value
F_Bigger_Minus_0.8 <- integrate(f_x, lower = -0.8, upper = 0)$value
answer <- F_Between0.8And0.6 / F_Bigger_Minus_0.8
paste("Question 2 answer is:", answer)
```

#### 3. Find the expectation of $X.$ [**10 marks**]
```{r}
# insert code
x.f_x <- function(x) x * 3 * (x^2)
E_X <- integrate(x.f_x, lower = -1, upper = 0)$value
paste("Question 3 answer is the expectation is:", E_X)
```

#### 4. Find the variance of $X.$ [**10 marks**]
```{r}
x2.f_x <- function(x) (x^2) * 3 * (x^2)
E_X2 <- integrate(x2.f_x, lower = -1, upper = 0)$value
Var_X <- E_X2 - E_X^2
paste("Question 4 answer is the variance is:", Var_X)
```

### Suppose $X \sim \mathcal{N}(\mu = 3, \sigma^2 = 10).$

#### 5. Find $\Pr(X < 4\ |\ X > 1).$ [**10 marks**]
$$ X < 4 \cap  X > 1 = 1 < X < 4  $$ 
```{r}
mu <- 3
sigma2 <- 10
sigma <- sqrt(sigma2)
x_between_1_and_4 <- pnorm(4, mu, sigma) - pnorm(1, mu, sigma)
x_bigger_1 <- pnorm(1, mu, sigma, lower.tail = FALSE)

answer <- x_between_1_and_4 / x_bigger_1
paste("Question 5 answer is:", answer)
```

#### 6. Find $x_1$ such that: $\Pr(X > x_1) = 0.13.$ [**10 marks**]
```{r}
qnorm(0.13, mu, sigma, lower.tail = FALSE)
```

### Statistical inference [**40 marks**]

The width of bolts of fabric is normally distributed with a mean of 950 mm and a standard deviation of 100 mm. 

Sample data of 500 bolt widths has been collected here: https://acaimo.github.io/teaching/statistical_programming/ca/bwidth.RData

Import the data and estimate the $95\%$ confidence interval for $\mu$ and $\sigma^2.$
```{r}
load(url("https://acaimo.github.io/teaching/statistical_programming/ca/bwidth.RData"))
sample <- bwidth
n <- length(sample)
# To calculate conf interval of the mean
xbar <- mean(sample)
se <- sd(sample) / sqrt(n)
alpha <- 0.05
L <- xbar - qt(1 - alpha / 2, n - 1) * se
U <- xbar + qt(1 - alpha / 2, n - 1) * se
print(paste("Conf interval for mean is:", L, "and", U))

# To calculate conf. interval of the variance 
s2 <- var(sample)
L <- ((n - 1) * s2) / qchisq(1 - alpha / 2, n - 1)  
U <- ((n - 1) * s2) / qchisq(alpha / 2, n - 1)
print(paste("Conf interval for variance is:", L, "and", U))

```
