    An experiment was performed to compare the effectiveness of 4 different 
    treatments on postoperative dental pain. The treatments were combinations of codeine
    and acupuncture. The Pain Data set is on the website. Download it and read it into
    R. It consists of 5 variables as follows.
    (a) Painlevel: baseline pain tolerance.
    (b) Codeine: indicator variable, 1=no codeine administered, 2=codeine adminis-
    tered.
    (c) Acupuncture: indicator variable, 1=no acupuncture administered, 2=acupunc-
    ture administered.
    (d) Relief: Pain relief from self-reporting.
    (e) Treat: factor variable with the following factor; Treat Code
    1=no codeine, no acupuncture
    2=codeine, no acupuncture
    3=no codeine, acupuncture
    4=codeine, acupuncture

    (2.) To perform a regression analysis on these data use lm(.) with the factor(.) 
    function and treat as the single predictor. This will automatically create the
    required dummy variables and impose the set to zero constraint to get a solution. 
    To do this you must declare the variable treat as a factor variable. 

    (3.) Get the parameter estimate table and write down what each parameter in the parameter
    estimate table is estimating in the context of the experiment performed.

    (3.) Create the dummy variables by hand for fitting the same model as that given 
    using factor(.).

    (4.) Fit a new model, allowing for an effect of codeine and the effect of acupuncture
    separately. Con you see any advantages/disadvantages to this model?
  
```{r}
current_path = rstudioapi::getActiveDocumentContext()$path
setwd(dirname(current_path))
source("anovatab.R")
```

```{r}
pain=read.table("pain.txt",header=T)
pain
```

#### Regression analysis

```{r}
treatF <- factor(pain$treat)
fit_pain=lm(Relief~treatF,data=pain)
anovatab(fit_pain)
summary(fit_pain)
by(pain$Relief,pain$treat,mean)

```
#### Confidence interval for each parameter:

```{r}
confint(fit_pain)
```

#### Creating dummy variables by myself:
```{r}
pain$t1=pain$t2=pain$t3=pain$t4=0
pain$t1[pain$treat==1]=1 
pain$t2[pain$treat==2]=1 
pain$t3[pain$treat==3]=1 
pain$t4[pain$treat==4]=1

```


#### Fitting model with dummy:  
```{r}
fit_d_pain <- lm(Relief~t2+t3+t4, data=pain)
anovatab(fit_d_pain)
summary(fit_d_pain)
confint(fit_d_pain)
```

  
```{r}
CodeineF <- factor(pain$Codeine)
AcuF <- factor(pain$Acupuncture)
fit_s_pain=lm(Relief~CodeineF+AcuF,data=pain)
anovatab(fit_s_pain)
summary(fit_s_pain)

```

```{r}
data.frame(
separated=predict(fit_s_pain, data.frame(CodeineF2=c(0,0,1,1), AcuF2=c(0,1,0,1))),

together=predict(fit_pain, data.frame(treatF2=c(0,0,0,1), treatF3=c(0,0,1,0), treatF4=c(0,1,0,0)))
)
```

