---
title: "ProblemSheet1a"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem Sheet 1a

# Measures of Location

## 1. Describe each of the following measures of location including their the pros and cons

## i Mean
To get the mean (or average), you have to sum all the elements of a dataset and divide by the number of elements on it. The mean can be represented by the following formula:
$$ \bar{x}=\frac{\sum_{i=1}^n x_i}{n} $$
An example could be used to calculate the mean of the salary hour from the Merit Market company editors
```{r}
editorsSalaries=c(12, 10,11, 12,9,13,12) 


mean(editorsSalaries) 
```
Pros:  
1. Easy to calculate (for small sets, you can do the maths in your head)  
2. Is well understood, you can probably say to most of people: I sleep 7 hours per day on average.


Cons:  
It's sensitive to extreme values. Look what happens with the mean salary from Merit Market editors when mister Merit Jr. becomes 18 and starts working on the department:
```{r}
meritJrGenerousSalary <- 90
editorsSalaries=c(12, 10,11, 12,9,13,12,meritJrGenerousSalary) 

mean(editorsSalaries)
```
  
Suddenly, you can't rely on the mean to answer questions like: how much me, Mr. Norman NoMerit could make per hour working on that market agency  

## ii Median  
Median is the value in the middle of a list of values. If the list has an even number of values, it's the mean between the two in the middle.  

Pros: not as sensitive as mean for extremes. Using the previos salaries example, median still a pretty good measure to answer how much mister Mr. Norman NoMerit would expect to make joining Merit Market editors group:
```{r}
editorsSalaries=c(12, 10,11, 12,9,13,12) 


median(editorsSalaries) 
```

```{r}
meritJrGenerousSalary <- 90
editorsSalaries=c(12, 10,11, 12,9,13,12,meritJrGenerousSalary) 

median(editorsSalaries)
```
  
  
Cons: The set must be ordered, you can't make a calculation up in your head that easily. Also, It doesn't use all the data set so, if your distribution is left or right skewed, it might be misleading used alone


## iii Variance
The variance is, how much the data spread around the mean. It's calculated with the following formula:
$$ \sigma^2=\frac{\sum_{i=1}^n (x_i-\bar{x})^2}{n-1}, $$
  
It's frequently square rooted to become the stantad deviation and be used on other calculations. 

Pros: It uses all data and it's basis to calculate other stable measures largely used on statistics.  
Cons: It's not that easy to interpret by using it alone and can be influenced by extreme values on the dataset. Good look saying on your sleep tracking phone app: the standard deviation of your sleep time is 1.23...  


## iv Skewness
Is the measure of symmetry on a given distribution. It can indicate whether a distribution is symmetric, left or right skewed. It's formula basically says: get how far the mean is from the mode, then divide it by the standard deviation: 
$$ sk_1=\frac{\bar{x}-Mode(x)}{\sigma}.$$
Zero/near zero means the distribution is symmetric, high positive or high negative values means the distribution is right or left skewed respectevely.  

Pros: Gives an overview of how your data is spread, can be used to determine how to work with you data also.  

Cons: sensitive to cons from the mode like more than one mode in the distribution which can hide a even or bi-modal distribution.   


# Counting

## 2. How many different combinations of 4 cards can be made for a 52 card deck.  
The order doesn't matter here so that will be:
$$ \frac{52!}{4! * (52 - 4)!}  -> \frac{52!}{4! * 48!} -> \frac{52 * 51 * 50 * 49}{4!} -> \frac{6497400}{24} -> 270725 $$

## 3. A bank issues bank cards with PINs consisting of 4 digits, each one {0,1,2,...,9}. How many unique PINs are there if

## i. Any 4-digit code can be used.
That is 10 possibilities for each digit so, it's `10*10*10*10` = 10.000 or 
$$ 10^4 $$

## ii. The digits must be different.
That is 10 possibilities for the first digit and one less for each subsequent `10*9*8*7` = 5040 or
$$ \frac{10!} {(10-4)!} -> \frac{10 * 9 * 8 * 7 * 6!} {6!} -> 10 * 9 * 8 * 7 -> 5040$$


## 4. In a lottery, each ticket has 5 one-digit numbers 0-9 which is not repeated on it.

## i You win if your ticket has the digits in any order. What are the total number of possible combinations?
$$ \frac{10!}{5! * (10 - 5)!}  -> \frac{10*9*8*7*6* 5!}{5! * 5!} -> \frac {10*9*8*7* 6} {5!} -> \frac{30240}{120} -> 252 $$

##ii You would win only if your ticket has the digits in the required order. What are the total number of combinations?  
$$ \frac{10!} {(10-5)!} -> \frac{10*9*8*7*6*5!} {5!} -> 10 * 9 * 8 * 7 * 6-> 30240 $$

  
  
## 5. How many different combinations of 6 cards can be made for a 52 card deck if

## i) order matters.
$$ \frac{52!}{(52 - 6)!}  -> \frac{52*51*50*49*48*47*46!}{46!}->52*51*50*49*48*47-> 14.658.134.400$$
## ii) order does not matter
$$ \frac{52!}{6! * (52 - 6)!}  -> \frac{52*51*50*49*48*47*46!}{6! * 46!}->\frac{52*51*50*49*48*47}{6!}-> 20358520$$
  
  
  
  
## 6. A poker hand consists of 7 cards:

## i) How many different hands are possible, if order does not matter, 
## ii) How many hands can be made with at least one king and one queen.

## 7. In a game of 5 card poker what are the number of different possible hands are there?

## a) A hand with a pair

## b) A hand with two pair

## c) A hand with Three of a kind

## d) A hand with a Flush (all the same suit)
