---
output:
  html_document: default
  pdf_document: default
---

**Problem Sheet 2B Probability Mass Distributions**

```{r setup, include=FALSE}
library("tidyr")
```

**1. There are 30 candy covered chocolates in a bag M&M's. There is a .1 probability that
that the candy is red. If X is the number of red M&M's in the bag.**  
i Give the binomial probability mass function for X.  
p = 0.1, q = 0.9, n = 30
$$ \binom{30}{k} 0.1^k * 0.9^{30-k} $$
ii Find the probability of less than 2 red M&Ms in the bag.  
$$ p(0) = \binom{30}{0} 0.1^0 * 0.9^{30-0} -> 
   \frac{30!}{(30-0)! * 0!} * 1 * 0.0423 = 0.0423 $$ 
```{r}
#Or, doing it in R step by step:
 choose(30, 0) * (0.1^0) * (0.9^(30-0))
```

```{r}
#source tutorial: http://www.r-tutor.com/elementary-statistics/probability-distributions/binomial-distribution
#Or, doing it in R leveraging the function that calculates it directly:
probOfZero <- dbinom(0, size=30, prob=0.1) 
probOfZero
```

$$ p(1) = \binom{30}{1} 0.1^1 * 0.9^{30-1} -> 
   \frac{30!}{(30-1)! * 1!} * 0.1 * 0.0471 = 0.1413 $$
```{r}
#Or, doing it in R step by step:
 choose(30, 1) * (0.1^1) * (0.9^(30-1))
```  
```{r}
#Or, doing it in R leveraging the function that calculates it directly:
probOfOne <- dbinom(1, size=30, prob=0.1) 
probOfOne
```

```{r}
#Probability of less than 2:

probOfZero + probOfOne
```

```{r}
#Or putting all together in a cumulative function:
pbinom(1, size=30, prob=0.1)
```
**2. A baby wakes on average 0.25 times every hour.**  
i If X is the number of times a baby wakes in an hour, give the poisson probability
mass function for X.  
λ = 0.25 
$$ \frac{0.25^X e^{-0.25}}{X!}$$
ii If X is the number of times a baby wakes in eight hour, give the poisson probability mass function for X.
Average wakes up in 8 hours: 0.25 * 8 = 2  
$$ \frac{2^X e^{-2}}{X!}$$
iii What is the probability that the baby does not wake during the 8 hours.  
$$ 1 - \frac{2^0 * e^{-2}}{0!} -> \frac{1 * 0.1353} {1} -> 0.1353$$
```{r}
#using the poison function in R: 
ppois(0, lambda=2) 
```
      
**3. Give the features of a**  
i Geometric Experiment.  
"You play until you win".  
2 outcomes: success or failure  
Independent  
p is constant  
X means how many observations till success  
  
ii Binomial Experiment.  
"You play x amount of times"  
2 outcomes: success or failure  
Independent  
p is constant  
X means how many success in n observations


iii Poisson Experiment.  
Events happening randomly over time  
Events occur independently  
Rate of occusrence is well defined average  
The random variable (or X) is the number of events in a interval  
    
      
**4. Every day a production line makes 100 computers of which 10% are defective. If X is
the number of defective computers in a day.**  
a) Give the binomial probability mass function for X.  
p=0.1, q=0.9, n=100
$$ \binom{100}{x} 0.1^x * 0.9^{100-x} $$
b) Find the probability that there is more than 2 computers defective in a day.  

$$ probMoreThanTwo = 1-p(2)-p(1)-p(0) $$ 
$$ 1 - \binom{100}{2} 0.1^2 * 0.9^{100-2} - \binom{100}{1} 0.1^1 * 0.9^{100-1} - \binom{100}{0} 0.1^0 * 0.9^{100-0}$$
```{r}
1 - pbinom(2, size=100, prob=0.1)
```
    
      
**5. A phone center receives 15 calls every 30 minutes.**  
a) If X is the number of phone calls in 30 minutes, give the poisson probability mass
function for X.  
$$ \frac{15^X e^{-15}}{X!}$$
b) What is the probability that there will be exactly 10 phone calls in the first 30
minutes and exactly 20 phone calls in the second 30 minutes.
$$ \frac{15^{10} e^{-15}}{10!} * \frac{15^{20} e^{-15}}{20!} $$
```{r}
dpois(10, lambda = 15) * dpois(10, lambda = 15)
```