---
output:
  html_document: default
  pdf_document: default
---

**Problem Sheet 2a Probability Mass Distributions**

```{r setup, include=FALSE}
library("tidyr")
```


1.  The probability mass function of a discrete random variable X is given in the following table:
```{r}
data.frame("x" = "p(x)", "0" = 0.1, "1"=0.2, "2"=0.4, "3"=0.2, "4"=0.1)
```
    
Find the E[X] and Var[X]  
k = 4, given the table above then:   
$$ E[X] = \sum_{k = 0}^{4} kp(k) -> 0*0.1 + 1*0.2 + 2*0.4 + 3*0.2 + 4*0.1 = 2 $$
$$ E[x] = n*p -> 2 = 4 * p  -> p = 0.5$$
$$ Var[x] = n*p*(1-p) -> 4*0.5 *(1 - 0.5) -> Var[x] = 1$$ //TODO WATCH 20:00 to 23:00
2.  The probability mass function of a discrete random variable X is given in the following table:

    **x | -2  | -1 | 0  | 1 | 2
    p(x)| 0.1 | 0.3| 0.3| p3| 0.1**
  
Show that p3 = 0.2 and calculate the E[X].

$$ p3 = 1 - (p(-2) + P(-1) + P(0) + P(2)) = 1 - (0.1 + 0.3 +  0.3 + 0.1) = 1 - 0.8 = 0.2  $$
$$ E[X] = \sum_{k = -2}^{2} kp(k) -> -2*0.1 - 1*0.3 + 0*0.3 + 1*0.2 + 2*0.1 = -0.1 $$
  
  
3. 20% of the Irish population watched Ireland beat France in the Rugby World Cup. A
    representative from TV3 marketing was sent to Grafton Street to ask passerbys their
    opinion of the match coverage. Let X denote the number of people need to be asked
    til the marketer successfully finds some who watched the game.  

i Give the Geometric probability mass function for X.
p=0.2  
q=1-0.2 -> q=0.8

$$p(X) = q^{(k-1)} * p  -> P(X) = 0.8^{(k-1)} * 0.2 $$

ii Find the probability that the marketer had to ask exactly 2 people.  
$$P(2) = 0.8^{(2-1)} * 0.2 = 0.16  $$

iii What is the E[X] and Var[X] of the distribution.  
$$E[x] = 1/p -> E[x] = 5  $$
$$Var[x] = q/p^2 -> 0.8/0.2^2 = 0.8/0.04 -> Var[x] = 20  $$
