---
output:
  html_document: default

---



### 1.a) Write the formula of the linear regression, explaining the terms used in the ex-pression.
$$\hat{y} = \beta_0 + x_1 \beta_1$$  
Where $\hat{y}$ is the predicted brain wieght value, $\beta_0$ is the intercept (325.5734) in this case, and $\beta_1$ is the head size multiplier, (0.26343) in this case. $x_1$ is a observed head size.
$$\hat{y} = 325.5734 + 0.26343x_1 $$ 
  
### 1.b) What is the predicted brain weight for a head of size 1280 c m3.
$$\hat{y} = 325.5734 + 0.26343 * 1280 = 662.7638$$ 
  
### 1.c) Calculate the 95% confidence intervals for the coefficients.

Conf int 95%
```{r}
  abs(qt(p = 0.05, df = 235))
```
$$\hat{\beta_0} \pm t_{235,\alpha/2} S_{\hat{\beta_0}} ->
  325.5734 \pm 1.651364 * 47.14085 -> [403.42, 247.73]$$

$$\hat{\beta_1} \pm t_{235,\alpha/2} S_{\hat{\beta_1}} ->
  0.26343 \pm 1.651364 * 0.01291 -> [0.2847491, 0.2421109]$$
  

### 1.d) How much of the variance in brain weight does the model explain and is it statistically significant.  

The model explains 63.93% of the variance of brain weight. p-value: < 2.2e-16 So it's statistically relevant.  

  
    
### 2.a) Calculate the $\beta_0$ and $\beta_1$ estimates.  
$SS_{x,x} =31486154$ $Cov(x,y)=8294377$  
$$\beta_1 =  \frac{SS_{x,x}}{Cov(x,y)}= \frac{31486154}{8294377} = 3.796084$$
  
$$ \beta_0 = \overline{y} - \beta_1 \overline{x} = 1283 - 3.796084 * 3634 = -12511.97 $$
  
### 2.b)  Write the formula of the linear regression, explaining the terms used in the ex-pression.  
$$\hat{y} = -12511.97 + 3.796084x_1 $$
  
### 2.c) Calculate the t-value for the $\beta_0$ and $\beta_1$ estimates.
$$(\hat{\beta_0})_{tvalue} = \frac{\hat{\beta_0}}{S_{\hat{\beta_0}}} ->
  \frac{-12511.97}{47.14085} = -265.4167$$

$$(\hat{\beta_1})_{tvalue} = \frac{\hat{\beta_1}}{S_{\hat{\beta_1}}} ->
  \frac{3.796084}{0.01291} = 294.0421$$
  
### 2.d) Using a 95% criteria determine if the Coefficients are significant.
Conf int 95%
```{r}
  abs(qt(p = 0.05, df = 235))
```
  

$$\hat{\beta_0} \pm t_{235,\alpha/2} S_{\hat{\beta_0}} ->
  -12511.97 \pm 1.651364 * 47.14085 -> [-12434.12, -12589.82]$$
  
  
$$\hat{\beta_1} \pm t_{235,\alpha/2} S_{\hat{\beta_1}} ->
  3.796084 \pm 1.651364 * 0.01291 -> [3.774765, 3.817403]$$
  
  
  
    
### 3.a) Write the formula of the linear regression, explaining the terms used in the expression.  

$$\hat{y} = \beta_0 + headSize_i \beta_1 + ageGroup_i\beta_2 +  sex_i \beta3 $$  
  
### 3.b) What is the predicted brain weight for a head of size $1280cm^3$, Age group 2, and female
$$\hat{y} = 464.56281 + 1280 * 0.24421 + 1 * -23.96845 +  1 * -22.54325 = 730.6399 $$ 
  
  
  
### 3.c) How much of the variance in brain weight does the model explain and is it statistically significant.
65.28% and is relevant 
  
  
### 3.d) Explain how the degrees of freedom are calculated for the F statistic.  
That's the number of observed values -2. Here we have 233 DF so we had 235 observed values


