"A motor insurance company is conducting an analysis of historic
customer retention data. The response variable is a binary 
indicator of whether that customer renewed their car insurance at
the end of their policy period or did not (1=renewed, 0=did not 
renew). Potential predictors were: the age group of the customer;
the type of cover (comprehensive or third party fire & theft) and
the engine size of the insured vehicle (coded A, B and C in order of 
increasing engine size). 

(a) Read the insurance.csv data into R.

(b) Fit a logistic regression model to these data with retained 
as the reponse variable and age, cover and engine_size as 
predictors. Which predictors are related to the response?

(c) Find the estimated odds ratio that a customer in the 20-29 
age group will renew over a customer in the 50-60 age group - all
other variables being equal.

(d) Discuss the evidence for the predictor ‘cover’ being related
to the response. 

(e) Predict the probability that a customer with the following 
values of the predictors would renew: age= 40-49, cover = 
comprehensive, engine size= A.

(f) State what type of customer the intercept is modelling in 
this case. Calculate the odds that such a customer will renew and
determine a 95% confidence interval for this odds."



```{r}
current_path = rstudioapi::getActiveDocumentContext()$path
setwd(dirname(current_path))
```
(a)Read data  

```{r}
insurance=read.csv('insurance.csv',header=T)
```
(b) Fit a logistic regression model to these data with retained 
as the reponse variable and age, cover and engine_size as 
predictors. Which predictors are related to the response?  

```{r}
fit=glm(retained~factor(age)+factor(cover)+factor(engine_size),family=binomial(),data=insurance)
summary(fit)
```
A: age50-60 and factor(cover)Comprehensive are related to the response  
  
(c) Find the estimated odds ratio that a customer in the 20-29 
age group will renew over a customer in the 50-60 age group - all
other variables being equal.
```{r}
age50_60logOddsRatio <- fit$coefficients["factor(age)50-60"] * -1 # log odds ratio == - \beta_3
age50_60oddsRatio <- exp(age50_60logOddsRatio)
age50_60oddsRatio
```
  
(d) Discuss the evidence for the predictor ‘cover’ being related
to the response. 
  
- With a p-value of 0.01695 and a z-value of 2.388, we can say that cover is related to the response on this model with comprehensive cover contributing positively, 0.31110, for renewing an insurance.

(e) Predict the probability that a customer with the following values of the predictors would renew: age= 40-49, cover = comprehensive, engine size= A.

$$ \eta = \beta_0 + \beta_1(0) + \beta_2(1) + \beta_2(0) + \beta_3(0) + \beta_4(0) + \beta_5(1) + \beta_6(0) + \beta_7(0)$$
$$ \eta = 1.04557 + 0.08053 + 0.31110 -> \eta = 1.4372 $$
$$ P = \frac{e^{1.4372}}{1+e^{1.4372}}$$
```{r}
exp(1.4372)/ (1 + exp(1.4372))
```

(f) State what type of customer the intercept is modelling in 
this case. Calculate the odds that such a customer will renew and
determine a 95% confidence interval for this odds."
  
A 20-29 y/o customer with cover |"3rd Party, F & T" with engine size A

$$\eta = \beta_0 -> \eta = 1.04557$$
$$ odds = e^{1.04557} -> odds = 2.84502 $$
$$ CI_{log odds} = \eta \pm Z(s.e.) -> 1.04557 \pm 1.96 * 0.21558 -> CI = [0.6230332,1.468107]$$
$$ CI_{odds} = e^{CI_{log odds}} -> [e^{0.6230332}, e^{1.468107}] -> [1.864575, 4.341010]$$
Or, with R:
```{r}
confint.default(fit)
```
```{r}
exp(confint.default(fit))
```
```{r}
nd=data.frame(age='20-29',
              cover='3rd Party, F & T',
              engine_size='A')
pred=predict(fit,newdata=nd,se.fit=T,interval='confidence')
predci=c(pred$fit-1.96*pred$se.fit,pred$fit+1.96*pred$se.fit)
exp(predci)
```

