---
output:
  html_document: default

---

### 1.a) Write the formula of the logistic regression, explaining the terms used in the ex-pression. 
- Log Odds:
$$\eta_i = \beta_0 + x_1 \beta_1 -> \eta_i = 5.69788 + (-0.10991) * x_1)$$  
Where $\beta_0$ is the intercept, $\beta_1$ is the estimator for yards and $x_1$ is the number of yards in an observation.  

- Odds: 
$$ e^{\eta_i} $$  


- Probability: 
$$ p_i = \frac{e^{\eta_i}} {1 + e^{\eta_i}} $$  

  
### 1. b) Calculate the z-value of the estimates  
- For $\beta_0$:  
$$ z = \frac{\hat{\beta_0} - \beta^0}{se(\hat{\beta_0)}} ->z = \frac{5.69788 - 0}{0.45110}$$  


```{r}
z <- 5.69788/0.45110
z
```
  
- For $\beta_1$:  
$$ z = \frac{\hat{\beta_1} - \beta^0}{se(\hat{\beta_1)}} ->z = \frac{-0.10991 - 0}{0.01058}$$
```{r}
z <- -0.10991/0.01058
z
```
  
### 1.c)What are the estimated odds that a 60 yard field goal will succeed  

$$ \eta_i = 5.69788 + (-0.10991) * 60) $$
  
```{r}
eta60 <- 5.69788 + (-0.10991) * 60
eta60
```
  
Odds: $$ e^{-0.89672} $$  
```{r}
odds <- exp(-0.89672)
print(paste("Odds for 60 yd: ", odds, ":1", sep = ""))
```
  
### 1. d) What is the probability that a 30 yard field goal will succeed, and what is probability it will fail.  

```{r}
eta30 <- 5.69788 + (-0.10991) * 30
odds <- exp(eta30)
odds
```
  
$$ p_{30} = \frac{e^{\eta_{30}}} {1 + e^{\eta_{30}}} -> \frac{11.02957} {1 + 11.02957}$$  
  
```{r}
prob30 <- 11.02957 / (1 + 11.02957)
print(paste("Prob. Success:",prob30))
print(paste("Prob. Fail:",1-prob30))
```
  
  
---

### 2.a Write the formula of the logistic regression, explaining the terms used in the ex-pression.  
- Log Odds:  

$$\eta_i = \beta_0 + x_1 \beta_1 + x_2 \beta_2 + x_3 \beta_3 -> 
\eta_i = -7.6096 + 1.2046 * x_1 + 1.4210 * x_2 + 3.1973 * x_3)$$  
  
Where $\beta_0$ is the intercept.  
$\beta_1$ is Aslevel estimate and $x_1$ is Aslevel observation.  
$\beta_2$ is Sblevel estimate and $x_2$ is Sblevel observation.    
$\beta_3$ is LineamentProx estimate and $x_3$ is observation (1 for presence and 0 for abscence)
  
- Probability:
$$ p_i = \frac{e^{\eta_i}} {1 + e^{\eta_i}} $$  
Given an calculation of Log Estimate $\eta_i$, this formula can calculate the probability of finding gold.  
  
### 2. b) Calculate the z-value of the estimates. (TODO discover what is exactly $\beta^0$ ) 
$$ z = \frac{\hat{\beta_{theEstimateThatYouWantTheZed}} - \beta^0}{se(\hat{\beta_{theEstimateThatYouWantTheZed})}}$$  
```{r}
calculateZ <- function(betaVal, seVal, estName) {
  print (paste(estName, betaVal / seVal))
}

beta0 <- -7.6096
beta1 <- 1.2046
beta2 <- 1.4210
beta3 <- 3.1973

seBeta0 <- 3.1661
seBeta1 <- 0.4899
seBeta2 <- 0.7301
seBeta3 <- 1.8911

calculateZ(beta0,seBeta0,"Beta0")
calculateZ(beta1,seBeta1,"Beta1")
calculateZ(beta2,seBeta2,"Beta2")
calculateZ(beta3,seBeta3,"Beta3")
  
```

### 2. c) What are the probability of finding a gold deposit given an As level of 3, Sblevel of 4 and no lineament.  

$$ \eta_i = -7.6096 + 1.2046 * 3 + 1.4210 * 4 + 3.1973 * 0) $$  
```{r}
logOdds <- -7.6096 + 1.2046 * 3 + 1.4210 * 4 + 3.1973 * 0
odds <- exp(logOdds)
probability <- odds / (1 + odds)
probability
```
  
### 2. d) Given a 95% rejection criteria with a z-criteria of (-1.96,196), are the estimates of the coefficents significant.  
- Beta2 1.9463 and Beta3 1.6907 aren't significant  

### extra. 95% conf interval for each beta (see teacher's notes):
$$\beta +- 1.96 SE (\beta) $$


  