---
output:
  html_document: default
  pdf_document: default
---
Problem Sheet 1b
Mathematical Probability**
```{r setup, include=FALSE}
library("tidyr")
```
1\. The events A and B are defined on a sample space. The following probabilities are known: Pr(A)=
2/3; Pr(B)= 1/6 and Pr(A ∩B)= 1/6

a) Are A and B mutually exclusive and independent?  
As Pr(A ∩B) is not equal to zero, they are not mutually exclusive.  
Also, as Pr(A ∩B) is not equal to P(A) * P (B), they are not independent
```{r}
prA <- 2/3
prB <- 1/6
prAintersecB <- 1/6
prA * prB == prAintersecB
```
b) Find the following additional probabilities:
  
i Pr(notA)  
```{r}
1 - 2/3
```

ii Pr(A ∪B)    
Pr(A ∪B) = P(A) + P(B) - Pr(A ∩B) so:
```{r}
prA + prB - prAintersecB
```


iii Pr(A\|B)  
$$\frac{Pr(A ∩B)}{Pr(B)} $$
```{r}
prAintersecB / prB
```
      
2\. The events A and B are defined on a sample space. The following probabilities are known: Pr(A)=
1/2 ; Pr(B)= 1⁄2 and Pr(A ∪B)= 5⁄6.
```{r comment="Exercise 2 setup"}
prA <- 1/2
prB <- 1/2
prAunionB <- 5/6
```
Find the following additional probabilities:  
i Pr(A)= Pr(notA)  
Pr(notA) = 1 - Pr(A) so
```{r}
prNotA <- 1 - prA 
prA == prNotA
```

ii Pr(A ∩ B)  
$$Pr(A ∪B) = Pr(A) + Pr(B) - Pr(A ∩B) -> $$
$$ -> - Pr(A ∩B) = Pr(A ∪B) - Pr(A) - Pr(B) $$
```{r}
prAintersecB <- (prAunionB - prA - prB) * -1
prAintersecB
```
iii Pr(A\|B)
$$ Pr(A|B) =  \frac{Pr(A ∩B)}{p(B)} $$
```{r}
prAgivenB <- prAintersecB / prB
prAgivenB
```  
  
3\. The events A and B are defined on a sample space. For each of the following conditions write down the defining probability statement in terms of A and B.

i The probability of all events except A  
Can be Pr(notA) or 1 - Pr(A)  

ii A and B are mutually exclusive  
Pr(A ∩B) = 0  

iii A and B are mutually exclusive and exhaustive  
Pr(A ∩B) = 0 and Pr(A ∪B) = 1  

iv The probability of A and/or B  
$$Pr(A ∪B) -> P(A) + P(B) - P(A ∩ B) $$
v The probability of A given B  
$$ Pr(A|B) =  \frac{Pr(A ∩B)}{p(B)} $$  


4\. A standard green and a standard red die are thrown at the same time.  
i Define the sample space for all the events.  

```{r}
greenDieFaces <- c(1,2,3,4,5,6)
redDieFaces <- c(1,2,3,4,5,6)

crossing(greenDieFaces, redDieFaces)
```

ii What is probability that the sum of the scores on the two dice is greater than 6.  
$$ \frac{numberOfCasesThatSumGreaterThanSix}{totalNumberOfCombinations} $$
```{r}
totalNumberOfCombinations <- length(greenDieFaces) * length(redDieFaces)
totalBiggerThanSix = 0
for(gFace in greenDieFaces) {
  for(rFace in redDieFaces) {
    if(gFace + rFace > 6) {
      totalBiggerThanSix = totalBiggerThanSix + 1 
    }
  }
}
totalBiggerThanSix / totalNumberOfCombinations
```
iii What is the probability that the green die is less than 5.  
4/6  
iv That is the probability that the sum of the two die is greater than six given that the green die is
less than 5.
```{r}
totalNumberOfCombinations <- length(greenDieFaces) * length(redDieFaces)
totalBiggerThanSix = 0
for(gFace in greenDieFaces) {
  for(rFace in redDieFaces) {
    if(gFace + rFace > 6 && gFace < 5) {
      totalBiggerThanSix = totalBiggerThanSix + 1 
    }
  }
}
totalBiggerThanSix / totalNumberOfCombinations
```

**Conditional Probability**

5\. 20% of the population owns a smart-phone and a tablet, and 60% of the population own a smart-
phone. What is the probability that a someone owns a tablet given that they own a smart-phone?  
Pr(os) = 0.6, Pr(os ∩ ot) = 0.2  
$$Pr(ot|os) = \frac{Pr(os ∩ ot)}{Pr(os)}$$
```{r}
prOs = 0.6
prOsintersecPrOt = 0.2
prOtGivenOs = prOsintersecPrOt / prOs
prOtGivenOs
```
6\. The probability that it is Friday and that a student is absent is 0.03. Since there are 5 school days in
a week, the probability that it is Friday is 0.2. What is the probability that a student is absent given
that today is Friday?
Pr(f) = 0.2, Pr(f ∩ a) = 0.03
$$Pr(a|f) = \frac{Pr(f ∩ a)}{Pr(f)}$$
```{r}
prF = 0.2
prFintersecPrA = 0.03
prAGivenF = prFintersecPrA / prF
prAGivenF
```
**Bayes Probability**

7\. In 2015, of the 200 million people registered to vote in America 4% say they will vote for Donald
Trump. Of these voters 50% are climate change sceptics, while only 8.3% of the other voters are
climate change sceptics.  
a) Find the probability that a voter selected is a climate change skeptic.
Given Trump T, Climate Skeptic C
From Trump voters: 
$$P(T ∩ C) = 0.04 * 0.5 = 0.02 $$  

From others: 
$$P(!T ∩ C) = 0.96 * 0.083 = 0.079 $$  

Total Skeptcs (Also prevalence/prob in total): 
$$P(C) = 0.099 $$ 


b) If a voter is selected and they are a climate change sceptic, determine the probability that they will vote for Trump.
 


$$ P(T|C) = \frac{P(T ∩ C)}{P(c)} = \frac{0.02}{0.099} $$
```{r}
(0.02 / 0.099)
```



8\. In 2015, of the 200 million people registered to vote in America 38% say they will vote for Donald
Trump. Of these voters 20% are birthers (people who questioned that President Obama was born in
the US), while only 1% of the other voters are birthers.
Given Trump T, Birther B
a) Find the probability that a voter selected is a birther.  
From Trump voters P(T ∩ B) : 0.38 * 0.2 = 0.076  
from other voters P(!T ∩ B): 0.62 * 0.01 = 0.0062  
Probability of Birthers P(B) = 0.0822  

b) If a voter is selected and they are a birther, determine the probability that they will vote for Trump.  

$$ P(T|B) =  \frac{P(T ∩ B)}{P(B)} = \frac{0.076}{0.0822}  $$
```{r}
(0.076 / 0.0822)
```

9\. In a school 30% of the students are male. Of these males 20% have long hair, while 90% of the female students have long hair.  
Given Male M and Long Hair L  
i Find the probability that a student selected has long hair.\
of male P(M ∩ L) -> 0.3 * 0.2 = 0.06  
of not male P(!M ∩ L) -> 0.7 * 0.9 = 0.63  
P(L) -> 0.06 + 0.063 = 0.69  

ii If a student is selected and they have long hair, determine the probability that the student is male.  

$$ P(M|L) = \frac{P(M ∩ L)}{P(L)} = \frac{0.06}{0.69}$$  
  

```{r}
0.06 / 0.69
```
