```{r}
# setwd as the folder where this script is in
current_path = rstudioapi::getActiveDocumentContext()$path
setwd(dirname(current_path))
```

```{r warning=FALSE}
# Setup that will be used across the rest of this notebook:
#Install and load kable extra to render tables in a nicer way
#install.packages("kableExtra")
library(kableExtra)
#install and load tidyverse to do dataset manipulations
#install.packages("tidyverse")
library(tidyverse)
source("anovatab.R")
# set a seed to gel always the same "random" results
set.seed(29011987)

# read carbonfibre cvs into a a dataframe
carbon_fibre <- read.csv("carbon_fibre.csv", header=T)
head(carbon_fibre) %>% 
  kbl(caption = "Table 1: Fist lines of Carbon Fibre Dataset") %>%
  kable_styling()

```
This dataset have a response, strenght, and two categorical predictors: temperature (high and low), and additive(a, b, c or d).  
Let's first start with a Main Effects Model to understand the relationship between those predictors and the response:


## Model 1, Main Effects Model
Our Main Effects Model will be the following:
$$ y_i = \beta_0 + \beta_1 \delta_{i,b} + \beta_2 \delta_{i,c} + \beta_3 \gamma_{i,low} + \varepsilon_i $$
Where $\delta_{i,b}$ will be 1 in case of using additive b, or zero otherwise, same for $\delta_{i,c}$ for additive c. If both $\delta$ are zero, then additive a is being used.  
Also $\gamma_{i,low}$ will be 1 in case a low temperature was used, otherwise, it will be 0.
We can start accessing the relevance of this model by checking if any of the predictors have an effect on strength.  
$H_0: \beta_1=\beta_2=\beta_3=0$
$H_\alpha: \beta_j\neq0$ for some j $\epsilon{1,2,3}$
```{r}
# Transform categorical variables into factor in the dataset
carbon_fibre$additive <- factor(carbon_fibre$additive)
carbon_fibre$temperature <- factor(carbon_fibre$temperature)
fit1=lm(strength~additive+temperature, data = carbon_fibre)
anovatab(fit1)  %>% 
  kbl(caption = "Table 2: Anovatab for Main Effects Model") %>%
  kable_styling()
```
Based on the anovatab above, with p-value 0.0005418, we reject $H_0$ and conclude that at least one predictor affects strength.  
We can also check the coeficcients of the model and see if they are significant:
```{r}
summary(fit1)[["coefficients"]] %>% 
  kbl(caption = "Table 3: Coeficients for Main Effects Model") %>%
  kable_styling()
```
Based on the P values we can see that all predictors are significant on the current model and they all reduce the strength.
An important detail is that this model assumes that temperature and additive don't iteract with each other. To consider that, we need to move to an Iteration model.  
  

## Model 2, Iteraction Model  
Our new model with iteration terms is the following:
$$ y_i = \beta_0 + \beta_1 \delta_{i,b} + \beta_2 \delta_{i,c} + \beta_3 \gamma_{i,low} + \beta_4 \delta_{i,b} \gamma_{i,low} + \beta_5 \delta_{i,c} \gamma_{i,low} + \varepsilon_i $$
  
Where this part $\beta_4 \delta_{i,b} \gamma_{i,low} + \beta_5 \delta_{i,c} \gamma_{i,low}$ is the iteration part of the model.  

```{r}
# Add iteration to the model 
fit2=update(fit1,.~.+additive:temperature)
drop1(fit2,test='F')

```
In this new model, we can see that the iteraction `additive:temperature` is significant (p-value 0.002844). Although we can see on the coefficients became somewhat problematic with $\beta_1,\beta_2,\beta_3$ with p-values > 0.05.
```{r}
summary(fit2)[["coefficients"]] %>% 
  kbl(caption = "Table 4: Coeficients for Iteraction Model") %>%
  kable_styling()
```

As the iteraction part works and its predictors are significantly higher in value, let's ignore for now the problematic $\beta_1,\beta_2,\beta_3$ and see a plot of this model.

```{r}
library(emmeans);library(ggplot2)

g=emmip(fit2,temperature ~ additive)
g+theme(axis.text=element_text(size=14),
        axis.title=element_text(size=16,face="bold"))
```
Here we can see that additive a has nearly identical good performance in high or low temperatures, whereas additive b has the best performance of them all in a high temperature and the worst performance in a low temperature. We can also see that additive c performs badly on low temperature and better on high temperature but not with extreme difference as additive b.

```{r}
emmeans(fit2,pairwise~additive:temperature,adjust='fdr')$contrasts %>%
kbl(caption = "Table 5: Pairwise Contrasts for Iteraction Model") %>%
  kable_styling()
```
On the pairwise contrast table we can see the statistic significance between different temperatures and additives. As seen in the chart, there is no difference between additive a with high or low temperatures.  
Also, most importantly to notice, there is no difference in strength between additive b with high temperature and additive a (either with high or low temperature), both pairs have a very high p-value of 0.496. In fact, there is not even evidence that b at high temperature is better than c at a high temperature (p-value 0.167)

## Conclusion  
We've concluded with our interaction model that all additives a, b, and c are equally good on high temperature. Additive a though, seems to perform also as good as the others in low temperature. There is big evidence also that additive b and c perform poorly in creating carbon fibre material strength in low temperatures
